Generalized Linear Mixed Models

### 2009-04-02 docs/inla/inla.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@169 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@170 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

<<thedata1,echo=true,fig=false>>=
library(MASS)
data(bacteria)
head(bacteria)
@

\column{0.5\textwidth}
\begin{itemize}
	\item $Y_{ij}$ is the bacteria presence in child $i$ at time $j$
	\item $X_{ij}$ is a vector of indicator variables for treatment
	\item $V_i$ is each child's relative risk
\end{itemize}
\end{columns}
\end{frame}

@@

\begin{frame}
\frametitle{First generation of approximations}
\begin{block}{Can we use least squares?}
\begin{itemize}
	\item With no random effects, model fitting is done with Iterated Reweighted Least Squares, or Fisher Scoring
	\item For Poisson data, pretend $Y_i \sim N(\lambda_i, \lambda_i)$ and use least squares.  Though you need to know the variance to do this.
	\item Choose a starting value, then iterate between calculating the variances and estimating the parameters.
\end{itemize}
\end{block}

@@

\begin{frame}
\frametitle{Example: Bacteria}

@@

\begin{frame}
\frametitle{Markov Chain Monte Carlo}

@@

@@

\begin{frame}
\frametitle{Example: }

@@

>>>>

<<thedata1,fig=false>>=
library(MASS)
data(bacteria)
bacterianew <- bacteria[,-(2:3)]
bacterianew$y = as.integer(bacteria$y=="y")
levels(bacterianew$trt) <- c("placebo", "drug", "drugplus")
bacteria=bacterianew

head(bacteria)
#$
@

\column{0.5\textwidth}
\begin{itemize}
\item the ID column denotes repeated measurements on each child.
	\item $Y_{ij}$ is the bacteria presence in child $i$ at time $j$
	\item $X_{ij}$ is a vector of indicator variables for treatment
	\item $V_i$ is each child's relative risk
\end{itemize}
\end{columns}
\end{frame}

@@

\begin{frame}
\frametitle{First generation of approximations}

\begin{itemize}
	\item With no random effects, model fitting is done with Iterated Reweighted Least Squares, or Fisher Scoring
	\item For Poisson data, pretend $Y_i \sim N(\lambda_i, \lambda_i)$ and use least squares.  Though you need to know the variance to do this.
	\item Choose a starting value, then iterate between calculating the variances and estimating the parameters.
\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{Example: Bacteria}
{\scriptsize
<<bacteriaPQL,fig=false,echo=true>>=
library(MASS)
pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)
summary(pql)
@
}

@@

\begin{frame}
\frametitle{Advantages/Disadvantages}

\begin{block}{Advantages}
\begin{itemize}
	\item Quick and easy to use
	\item It's better than ignoring dependence completely and using a glm.
\end{itemize}
\end{block}

\begin{block}{Disadvantages}
\begin{itemize}
	\item It doesn't always produce good results
	\item Especially if your data are very non-normal,  with small numbers in each group.
	\item Theoretical justification isn't strong
\end{itemize}
\end{block}

@@

\begin{frame}
\frametitle{Markov Chain Monte Carlo}
\begin{itemize}
	\item Put prior distributions on parameters $[\sigma]$, $[\beta]$.
	\item Simulate from the posterior distributions
	\[
	[\sigma|Y],\ [\beta|Y],\ [V|Y]
	\]
	\item 
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Example: }
<<bacteriaINlA>>=
# do this pql = glmmPQL(y ~ trt, random = ~1|ID, family='binomial',data=bacteria)

@

@@

\begin{frame}
\frametitle{Example: Lung cancer in Ontario}
<<cancerInla>>=
library(diseasemapping)
library(sp)
data(popdata)
data(casedata)
model = getRates(casedata, popdata, ~age*sex)
ontario = getSMR(popdata,model,  casedata)
data(popDataAdjMat)

# fit bym model 
#forBugs = glmmBUGS(formula=Observed + logExpected ~ 1,
#  effects="CSDUID",   family="poisson", spatial=popDataAdjMat,
#  data=ontario)

@
\end{frame}
